Bayesian hidden Markov model analysis of single-molecule force spectroscopy:\\ Characterizing kinetics under measurement uncertainty

### 2015-05-18 manuscripts/bayesian-hmm-method/bhmm.tex
Revised synthetic model, but convergence seems to be poor.

\begin{figure*}[tbp]
\noindent
%\resizebox{\textwidth}{!}{\includegraphics{model-mlhmm-stateassignments}}
\resizebox{\textwidth}{!}{\includegraphics{synthetic-three-state-model-mlhmm}}
\caption{{\bf Synthetic force trajectory and inferred state assignments in MLHMM.}
Observed samples are colored by their hidden state assignments.
Dark horizontal lines terminating in triangles to the right denote state means, while lightly colored bands indicate one standard deviation on either side of the state mean.
The histograms on the right side shows the state-resolved probabilities of samples, while the colored peaks show the weighted Gaussian output contribution from each state, and the black outline the weighted sum of the Gaussian output contributions from the HMM states.
\label{figure:model-stateassignments}}
\end{figure*}

@@

\caption{{\bf Estimated mean model parameters and confidence intervals for synthetic timeseries data}}

@@

		Equilibrium probability 		& $\pi_{1}$ & $0.333$ & $0.338_{\:0.130}^{\:0.543}$  & $0.309_{\:0.251}^{\:0.374}$  & $0.330_{\:0.299}^{\:0.363}$  \\
		& $\pi_{2}$ & $0.333$ & $0.351_{\:0.082}^{\:0.735}$  & $0.335_{\:0.246}^{\:0.432}$  & $0.329_{\:0.309}^{\:0.350}$  \\
		& $\pi_{3}$ & $0.333$ & $0.312_{\:0.118}^{\:0.510}$  & $0.356_{\:0.299}^{\:0.411}$  & $0.341_{\:0.322}^{\:0.357}$  \\

@@

		Transition probability ($\Delta t = $1.0 ms) 		& $T_{11}$ & $0.989$ & $0.952_{\:0.926}^{\:0.972}$ & $0.947_{\:0.939}^{\:0.955}$ & $0.988_{\:0.987}^{\:0.989}$ \\
		& $T_{12}$ & $0.010$ & $0.005_{\:0.001}^{\:0.015}$ & $0.001_{\:0.001}^{\:0.003}$ & $0.001_{\:0.001}^{\:0.001}$ \\
		& $T_{13}$ & $0.001$ & $0.043_{\:0.024}^{\:0.067}$ & $0.051_{\:0.044}^{\:0.060}$ & $0.011_{\:0.010}^{\:0.012}$ \\
		& $T_{21}$ & $0.010$ & $0.005_{\:0.001}^{\:0.016}$ & $0.001_{\:0.000}^{\:0.003}$ & $0.001_{\:0.001}^{\:0.001}$ \\
		& $T_{22}$ & $0.940$ & $0.990_{\:0.975}^{\:0.998}$ & $0.988_{\:0.983}^{\:0.991}$ & $0.949_{\:0.946}^{\:0.951}$ \\
		& $T_{23}$ & $0.050$ & $0.005_{\:0.001}^{\:0.015}$ & $0.011_{\:0.008}^{\:0.015}$ & $0.050_{\:0.048}^{\:0.053}$ \\
		& $T_{31}$ & $0.001$ & $0.046_{\:0.028}^{\:0.072}$ & $0.044_{\:0.038}^{\:0.052}$ & $0.011_{\:0.009}^{\:0.012}$ \\
		& $T_{32}$ & $0.050$ & $0.005_{\:0.001}^{\:0.016}$ & $0.010_{\:0.007}^{\:0.014}$ & $0.049_{\:0.046}^{\:0.051}$ \\
		& $T_{33}$ & $0.949$ & $0.949_{\:0.922}^{\:0.970}$ & $0.945_{\:0.937}^{\:0.953}$ & $0.941_{\:0.938}^{\:0.943}$ \\

@@

		Transition rate (ms$^{-1}$) 		& $k_{12}$ & $0.010$ & $0.005_{\:0.001}^{\:0.015}$ & $0.001_{\:0.001}^{\:0.003}$ & $0.001_{\:0.001}^{\:0.001}$ \\
		& $k_{13}$ & $0.001$ & $0.043_{\:0.024}^{\:0.067}$ & $0.051_{\:0.044}^{\:0.060}$ & $0.011_{\:0.010}^{\:0.012}$ \\
		& $k_{21}$ & $0.010$ & $0.005_{\:0.001}^{\:0.016}$ & $0.001_{\:0.000}^{\:0.003}$ & $0.001_{\:0.001}^{\:0.001}$ \\
		& $k_{23}$ & $0.050$ & $0.005_{\:0.001}^{\:0.015}$ & $0.011_{\:0.008}^{\:0.015}$ & $0.050_{\:0.048}^{\:0.053}$ \\
		& $k_{31}$ & $0.001$ & $0.046_{\:0.028}^{\:0.072}$ & $0.044_{\:0.038}^{\:0.052}$ & $0.011_{\:0.009}^{\:0.012}$ \\
		& $k_{32}$ & $0.050$ & $0.005_{\:0.001}^{\:0.016}$ & $0.010_{\:0.007}^{\:0.014}$ & $0.049_{\:0.046}^{\:0.051}$ \\

@@

		State mean lifetime (ms) 		& $t_{1}$ & $90.408$ & $21.797_{\:13.184}^{\:36.162}$ & $18.645_{\:15.899}^{\:21.524}$ & $83.405_{\:75.423}^{\:92.188}$ \\
		& $t_{2}$ & $16.162$ & $148.113_{\:43.004}^{\:525.609}$ & $82.451_{\:59.320}^{\:115.043}$ & $18.954_{\:18.001}^{\:19.919}$ \\
		& $t_{3}$ & $19.103$ & $20.138_{\:12.475}^{\:33.395}$ & $17.902_{\:15.417}^{\:20.637}$ & $16.395_{\:15.692}^{\:17.177}$ \\

@@

		Relaxation time (ms) 		& $\tau_{1}$ & $62.700$ & $82.300_{\:32.985}^{\:200.069}$ & $56.416_{\:44.005}^{\:70.295}$ & $58.230_{\:54.026}^{\:62.826}$ \\
		& $\tau_{2}$ & $8.909$ & $10.445_{\:7.396}^{\:15.490}$ & $9.353_{\:8.387}^{\:10.399}$ & $8.970_{\:8.681}^{\:9.287}$ \\

@@

		State force mean (pN) 		& $\mu_{1}$ & $3.000$ & $4.677_{\:4.644}^{\:4.712}$ & $5.606_{\:5.600}^{\:5.613}$ & $3.002_{\:2.991}^{\:3.013}$ \\
		& $\mu_{2}$ & $4.700$ & $3.069_{\:2.953}^{\:3.187}$ & $3.019_{\:2.986}^{\:3.056}$ & $5.602_{\:5.600}^{\:5.604}$ \\
		& $\mu_{3}$ & $5.600$ & $5.602_{\:5.580}^{\:5.625}$ & $4.700_{\:4.690}^{\:4.710}$ & $4.703_{\:4.699}^{\:4.706}$ \\

@@

		State force std dev (pN) 		& $s_{1}$ & $1.000$ & $0.311_{\:0.285}^{\:0.337}$ & $0.199_{\:0.194}^{\:0.203}$ & $1.002_{\:0.995}^{\:1.010}$ \\
		& $s_{2}$ & $0.300$ & $1.067_{\:0.992}^{\:1.147}$ & $1.012_{\:0.988}^{\:1.039}$ & $0.200_{\:0.199}^{\:0.202}$ \\
		& $s_{3}$ & $0.200$ & $0.210_{\:0.195}^{\:0.229}$ & $0.297_{\:0.290}^{\:0.305}$ & $0.299_{\:0.296}^{\:0.301}$ \\

>>>>

\begin{figure*}[tbp]
\noindent
%\resizebox{\textwidth}{!}{\includegraphics{model-mlhmm-stateassignments}}
%\resizebox{\textwidth}{!}{\includegraphics{synthetic-three-state-model-mlhmm}}
\resizebox{\textwidth}{!}{\includegraphics{synthetic-three-state-model-mlhmm-nstates3-nobs10000}}
\caption{{\bf Synthetic force trajectory and inferred state assignments in MLHMM.}
Observed samples are colored by their hidden state assignments.
Dark horizontal lines terminating in triangles to the right denote state means, while lightly colored bands indicate one standard deviation on either side of the state mean.
The histograms on the right side shows the state-resolved probabilities of samples, while the colored peaks show the weighted Gaussian output contribution from each state, and the black outline the weighted sum of the Gaussian output contributions from the HMM states.
\label{figure:model-stateassignments}}
\end{figure*}

@@

\caption{{\bf Estimated mean model parameters and confidence intervals for synthetic timeseries data (new table)}}

@@

		Equilibrium probability 		& $\pi_{1}$ & $0.385$ & $0.678_{\:0.367}^{\:0.889}$  & $0.533_{\:0.442}^{\:0.629}$  & $0.517_{\:0.486}^{\:0.549}$  \\
		& $\pi_{2}$ & $0.094$ & $0.246_{\:0.069}^{\:0.546}$  & $0.351_{\:0.264}^{\:0.434}$  & $0.381_{\:0.353}^{\:0.410}$  \\
		& $\pi_{3}$ & $0.521$ & $0.077_{\:0.029}^{\:0.156}$  & $0.116_{\:0.091}^{\:0.144}$  & $0.101_{\:0.093}^{\:0.110}$  \\

@@

		Transition probability ($\Delta t = $1.0 ms) 		& $T_{11}$ & $0.980$ & $0.991_{\:0.981}^{\:0.997}$ & $0.989_{\:0.986}^{\:0.992}$ & $0.989_{\:0.988}^{\:0.990}$ \\
		& $T_{12}$ & $0.015$ & $0.001_{\:0.000}^{\:0.006}$ & $0.003_{\:0.002}^{\:0.005}$ & $0.004_{\:0.003}^{\:0.004}$ \\
		& $T_{13}$ & $0.005$ & $0.009_{\:0.003}^{\:0.018}$ & $0.008_{\:0.006}^{\:0.010}$ & $0.007_{\:0.006}^{\:0.008}$ \\
		& $T_{21}$ & $0.063$ & $0.003_{\:0.000}^{\:0.017}$ & $0.005_{\:0.003}^{\:0.008}$ & $0.005_{\:0.004}^{\:0.005}$ \\
		& $T_{22}$ & $0.900$ & $0.967_{\:0.932}^{\:0.988}$ & $0.976_{\:0.970}^{\:0.982}$ & $0.978_{\:0.977}^{\:0.980}$ \\
		& $T_{23}$ & $0.037$ & $0.030_{\:0.010}^{\:0.068}$ & $0.019_{\:0.014}^{\:0.025}$ & $0.017_{\:0.015}^{\:0.018}$ \\
		& $T_{31}$ & $0.003$ & $0.079_{\:0.029}^{\:0.146}$ & $0.036_{\:0.026}^{\:0.047}$ & $0.036_{\:0.033}^{\:0.040}$ \\
		& $T_{32}$ & $0.007$ & $0.093_{\:0.028}^{\:0.193}$ & $0.057_{\:0.043}^{\:0.072}$ & $0.063_{\:0.058}^{\:0.068}$ \\
		& $T_{33}$ & $0.990$ & $0.828_{\:0.711}^{\:0.920}$ & $0.907_{\:0.887}^{\:0.925}$ & $0.901_{\:0.894}^{\:0.907}$ \\

@@

		Transition rate (ms$^{-1}$) 		& $k_{12}$ & $0.015$ & $0.001_{\:0.000}^{\:0.006}$ & $0.003_{\:0.002}^{\:0.005}$ & $0.004_{\:0.003}^{\:0.004}$ \\
		& $k_{13}$ & $0.005$ & $0.009_{\:0.003}^{\:0.018}$ & $0.008_{\:0.006}^{\:0.010}$ & $0.007_{\:0.006}^{\:0.008}$ \\
		& $k_{21}$ & $0.063$ & $0.003_{\:0.000}^{\:0.017}$ & $0.005_{\:0.003}^{\:0.008}$ & $0.005_{\:0.004}^{\:0.005}$ \\
		& $k_{23}$ & $0.037$ & $0.030_{\:0.010}^{\:0.068}$ & $0.019_{\:0.014}^{\:0.025}$ & $0.017_{\:0.015}^{\:0.018}$ \\
		& $k_{31}$ & $0.003$ & $0.079_{\:0.029}^{\:0.146}$ & $0.036_{\:0.026}^{\:0.047}$ & $0.036_{\:0.033}^{\:0.040}$ \\
		& $k_{32}$ & $0.007$ & $0.093_{\:0.028}^{\:0.193}$ & $0.057_{\:0.043}^{\:0.072}$ & $0.063_{\:0.058}^{\:0.068}$ \\

@@

		State mean lifetime (ms) 		& $t_{1}$ & $49.498$ & $128.116_{\:55.334}^{\:309.568}$ & $91.902_{\:70.769}^{\:120.147}$ & $93.429_{\:86.014}^{\:101.962}$ \\
		& $t_{2}$ & $9.491$ & $36.448_{\:15.042}^{\:92.972}$ & $42.545_{\:33.175}^{\:54.544}$ & $45.889_{\:42.634}^{\:49.415}$ \\
		& $t_{3}$ & $99.499$ & $5.986_{\:3.065}^{\:13.200}$ & $10.396_{\:8.396}^{\:12.904}$ & $9.555_{\:8.969}^{\:10.219}$ \\

@@

		Relaxation time (ms) 		& $\tau_{1}$ & $56.704$ & $54.467_{\:26.737}^{\:109.929}$ & $51.201_{\:41.887}^{\:62.042}$ & $53.387_{\:50.301}^{\:56.513}$ \\
		& $\tau_{2}$ & $8.377$ & $5.191_{\:2.656}^{\:10.780}$ & $8.855_{\:7.255}^{\:10.815}$ & $8.335_{\:7.857}^{\:8.866}$ \\

@@

		State force mean (pN) 		& $\mu_{1}$ & $3.000$ & $5.603_{\:5.588}^{\:5.619}$ & $5.600_{\:5.595}^{\:5.606}$ & $5.599_{\:5.597}^{\:5.600}$ \\
		& $\mu_{2}$ & $4.700$ & $2.950_{\:2.818}^{\:3.085}$ & $3.003_{\:2.970}^{\:3.039}$ & $2.996_{\:2.986}^{\:3.007}$ \\
		& $\mu_{3}$ & $5.600$ & $4.706_{\:4.598}^{\:4.822}$ & $4.712_{\:4.694}^{\:4.731}$ & $4.700_{\:4.694}^{\:4.706}$ \\

@@

		State force std dev (pN) 		& $s_{1}$ & $1.000$ & $0.198_{\:0.188}^{\:0.209}$ & $0.198_{\:0.195}^{\:0.202}$ & $0.200_{\:0.198}^{\:0.201}$ \\
		& $s_{2}$ & $0.300$ & $1.011_{\:0.918}^{\:1.114}$ & $0.994_{\:0.970}^{\:1.018}$ & $0.998_{\:0.991}^{\:1.005}$ \\
		& $s_{3}$ & $0.200$ & $0.417_{\:0.290}^{\:0.524}$ & $0.291_{\:0.277}^{\:0.305}$ & $0.298_{\:0.294}^{\:0.303}$ \\
